Moral Foundations of U.S. Political News Organizations

### 2019-04-24 pilot_manuscript.tex
Fixing shit for the knit (it rhymes)

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

@@

\abstract{
The media ecosystem has grown, and political opinions have diverged such
that there are competing conceptions of objective truth. Commentators
often point to political biases in news coverage as a catalyst for this
political divide. The Moral Foundations Dictionary (MFD) facilitates
identification of ideological leanings in text through frequency of the
occurrence of certain words. Through web scraping, the researchers
extracted articles from popular news sources' websites, calculated MFD
word frequencies, and identified words' respective valences. This
process attempts to uncover news outlets' positive or negative
endorsements of certain moral dimensions concomitant with a particular
ideology. In Experiment 1, the researchers gathered poltical articles
from four sources. They were unable to reveal significant differences in
moral or political endorsements, but they solidified the method to be
employed in further research. In Experiment 2, the researchers expanded
their number of sources to 10 and analyzed articles that pertain to two
specific topics: the 2018 confirmation hearings of U.S. Supreme Court
Justice Brett Kavanaugh and the partial U.S. Goverment Shutdown of
2018-2019. Once again, no significant differences in moral or political
endorsements were found.

@@

Moral Foundations Theory has received criticism on the grounds that its
assumptions regarding moral intuitions have little empirical basis.
Suhler and Churchland (2011) list several potential weaknesses of MFT
that they argue might threaten the theory's validity. First, the authors
challenge Haidt and Graham (2007)'s claims the moral intuitions
represented by MFT are innate and modular. Suhler and Churchland (2011)
claim that advances in biological sciences (embryology and microbiology,
specifically) make it more difficult for researchers to claim any one
trait is either innate or learned through experience. Rather, behaviors
likely result from interactions between genetics and experience (Suhler
\& Churchland, 2011). According to the authors, without solid data
supporting the innateness of moral foundations, Haidt and Graham have
little from which to make such a claim. Similarly, Haidt and Graham
(2007) rely on evidence authored by evolutionary psychologists to make a
\enquote{strong modularity claim} (Suhler \& Churchland, 2011). However,
as with innateness, there is little neurobiological evidence to support
modularity.

Suhler and Churchland (2011) also criticized the content and taxonomy of
the five foundations. The authors criticize Haidt and Graham (2007)'s
omissions of additional foundations, including \emph{industry} and
\emph{modesty}, claiming these concepts are moralized in many societies
worldwide. Likewise, the authors question whether or not the foundations
are sufficiently distinct as to stand as their own foundation. For
example, Suhler and Churchland (2011) posit that \emph{ingroup/loyalty}
is merely a group-focused version of \emph{harm/care}. Work by Graham et
al. (2011) might serve to rebut this criticism, as the researchers found
their original five-factor structure seemed to best fit the data when
validating the Moral Foundations Questionnaire. Finally, Suhler and
Churchland (2011) point out that particular concepts related to a
foundation, including \enquote{anger} in \emph{fairness/reciprocity} and
\enquote{deception} in \emph{ingroup/loyalty}, could be ascribed to any
of the other foundations as well. In other words, it becomes difficult
to recognize a particular concept as indicative of any one foundation
when, in theory, they could be applied to all five (Haidt \& Graham,
2007; Suhler \& Churchland, 2011).

These criticisms of Moral Foundations Theory are valid and should be
taken into consideration when conducting research with instruments
derived from MFT. The current authors argue these criticisms are
especially valid when considered alongside questions regarding the Moral
Foundations Dictionary they state herein. However, there exists
compelling evidence regarding the validity of Moral Foundations Theory,
albeit regarding its application solely through the Moral Foundations
Questionnaire. This evidence is discussed within a brief explanation of
the Moral Foundations Questionnaire and its relationship to the Moral
Foundations Dictionary.

@@

In order to capture language's role in moral and political reasoning,
Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD)
in order to capture moral reasoning and justification as used in speech
and text. The MFD is composed of 259 words, with around 50 words
assigned to each of the five foundations. The researchers created a
preliminary list of words that they believed would be associated with
the five foundations. Then, using the Linguistic Inquiry and Word Count
(LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they
analyzed transcripts of liberal and conservative Christian sermons in
order to obtain frequencies of the occurrence of words from the
researchers' initial list. The researchers manually checked the results
from LIWC in order to make sure the results make sense given the
contexts and rhetorical devices used in the sermons, as word frequency
analysis ignores sentence context. The researchers offered the following
example from a Unitarian sermon as a demonstration of ambiguous
statements requiring human verification: \enquote{Don't let some
self-interested ecclesiastical or government authority tell you what to
believe, but read the Bible with your own eyes and open your heart
directly to Jesus} (Graham et al., 2009). This sentence added to the
\emph{authority/respect} total in LIWC's analysis, but it appears to
suggest that one should reject authority in this context. The
researchers eliminated this sentence from the \emph{authority/respect}
raw count on account of this discrepancy between the use of
authority-related words and the speaker's clear intentions (Graham et
al., 2009).

@@

The Moral Foundations Questionnaire has been validated by multiple
researchers. Likewise, its five-factor structure has been demonstrated
to fit data in multiple countries (Davies, Sibley, \& Liu, 2014; Graham
et al., 2011). In their article introducing the 30-item MFQ, Graham et
al. (2011) conducted an exploratory factor analysis (EFA) and found that
a five-factor model fit the data better than a one, two, or three-factor
model. Davies et al. (2014) conducted a confirmatory factor analysis
(CFA) of the MFQ with a sample from New Zealand and likewise found the
five-factor model provided the best fit. While Davies et al. (2014)
concede the US and New Zealand share many similarities as Western
nations, which could raise questions regarding the validity of the MFQ
in non-Western nations. However, there are striking differences between
the two countries, including the lack of a two-party political system in
New Zealand, that provide grounds for claiming the MFQ generalizes
beyond the United States. Furthermore, Graham et al. (2011) claim to
find a \enquote{reasonable} degree of generalizability for the MFQ
across participants from many different regions in the world. These two
bodies of work also provide the best available evidence that the five
moral foundations are sufficiently distinct from one another, though
broader criticisms of MFT raised by Suhler and Churchland (2011) should
still be taken into account in studies involving the theory.

@@

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
sources (identified using CSS language) and compiled them into a dataset
(Wickham, 2016). Using this dataset, we identified word count and
average word count per source. This process was run once daily starting
in February 2018 until March 2018. Starting in mid-March 2018, the
process was run twice daily - once in the morning and again in the
evening. Data collection was terminated once 250,000 words per source
was collected in April 2018.

@@

As seen in Table \ref{tab:exp1-source-descriptives}, the sources are
similar in some aspects yet different in others. Valence appears to be
slightly positive across all sources. The large standard deviations seem
to indicate little to no presence of a difference in valence. \emph{The
New York Times} published the greatest number of articles as well as
total words. \emph{Breitbart} featured the lowest number of articles,
and \emph{NPR} the lowest number of total words from all articles. Per
individual article, however, \emph{Breitbart} appears to feature the
highest average number of words as well as unique words. Once again the
standard deviations call into question any apparent differences between
sources. Finally, \emph{Fox News} appears to be the most readable source
on average, while \emph{The New York Times} is the least readable. This
result might be attributable to the greater number of tokens in the
average \emph{New York Times} article compared to \emph{Fox News}. As
before, the standard deviations for readability hamper one's ability to
draw solid conclusions regarding differences in the sources' use of
language.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp1-table}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .850 & 0.01\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .715 & 0.02\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .789 & -0.04\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.17 & .087 & -0.20\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.37 & .141 & -0.09\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

@@

Table \ref{tab:exp2-source-descriptives-kav} displays the descriptive
statistics for sources' writing on the Kavanaugh confirmation hearing.
The sources were similar in most basic linguistic aspects, except for
number of articles. For example, \emph{Sean hannity} appears to have
published only 27 articles while \emph{Breitbart} published 757 articles
on this topic. Valence was found to be slightly positive across all
sources. \emph{Fox News} produced the most total words with the most
tokens on average. \emph{Politico} featured the highest number of types
on average. \emph{Rush Limbaugh} was the most readable source on average
by grade level while \emph{Slate} was the least readable. The large
standard deviations for these statistics, however, preclude conclusions
regarding differences in the sources' use of language.

Table \ref{tab:exp2-source-descriptives-gs} displays descriptive
statistics for articles about the partial government shutdown. Like the
Kavanaugh hearing, the sources were similar in average valence (slightly
positive). Once again, there was variation in the number of articles
published by each source on this topic. \emph{Sean Hannity}, \emph{Rush
Limbaugh}, and \emph{The Blaze} published fewer than 100 articles while
\emph{Fox News} published 1,013 articles. \emph{Fox News} again featured
the most total words and mean tokens. \emph{Politico} had the most types
on average. For this topic, \emph{Fox News} was the most readable while
\emph{Slate} was the least readable. For each statistic, the excessively
high standard deviations render any assertions regarding linguistic
differences inconclusive on a descriptive level.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablekav}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.39 & 1.35 & 0.66 & 1.35 & 2.89 & .020 & -0.19\\
Fairness/Reciprocity & 1.26 & 1.27 & 1.06 & 1.27 & -1.58 & .154 & 0.16\\
Ingroup/Loyalty & 1.15 & 1.22 & 1.06 & 1.22 & -0.63 & .549 & 0.07\\
Authority/Respect & 0.85 & 0.99 & 0.75 & 0.99 & -0.93 & .380 & 0.10\\
Purity/Sanctity & 0.90 & 1.28 & 1.06 & 1.28 & 0.88 & .406 & -0.11\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablegs}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.91 & 1.27 & 1.01 & 1.27 & 1.68 & .132 & -0.07\\
Fairness/Reciprocity & 1.22 & 1.05 & 0.91 & 1.05 & -0.76 & .468 & 0.27\\
Ingroup/Loyalty & 1.38 & 1.16 & 1.11 & 1.16 & -0.48 & .644 & 0.21\\
Authority/Respect & 0.87 & 1.24 & 0.36 & 1.24 & -1.38 & .203 & 0.37\\
Purity/Sanctity & 1.29 & 1.13 & 0.99 & 1.13 & 0.03 & .979 & 0.22\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

@@

The results obtained in both experiments, while not confirmatory,
provide several avenues for future investigation. Twice confronted with
the failure to uncover differences in news media's political content,
the researchers were unable to confirm the existence of an effect of
political lean on American news writing under the auspices of Moral
Foundations Theory. This does not, however, rule out the possibility
that news media is biased. As Mitchell et al. (2014) has demonstrated,
self-reported political partisans are drawn to separate sources, but the
present research was unable to offer any content-based explanations of
this phenomenon. Even when attenuating the scraping criteria to collect
text concerning (what the researchers believed to be) morally resonant
topics, the method failed to detect differences in MFD word use by
prominent news sources. The researchers' focus, therefore, turns toward
the limitations of the method, especially the chosen theoretical
framework, under which the present research was conducted.

@@

Thus, the researchers' attention turned to the instrument used to detect
differences. In the present study, the researchers utilized the Moral
Foundations Dictionary, a metric developed from Moral Foundations Theory
(Graham et al., 2009). Based upon the results obtained, it might be
necessary to investigate alternative instruments that might better
elucidate the differences of interest. Likewise, other theoretical
perspectives may be better equipped to explain political differences in
discourse.

Tools such as these could be developed through the Linguistic Inquiry
and Word Count program (LIWC) (Pennebaker et al., 2007). As stated
before, the Moral Foundations Dictionary was established through an
analysis of Christian sermons performed mostly with LIWC as well as some
human-enforced corrections due to context (Graham et al., 2009). This
approach was grounded in Moral Foundations Theory, a clear framework
within which the authors developed the MFD. The original authors of the
MFD established sound psychometric properties for the tool, implying its
validity in several applications. For this reason, the researchers
elected to utilize the MFD as their primary instrument.

Other instruments could theoretically be developed through a LIWC
analysis, including those that would be useful for analyzing political
discourse. Researchers could obtain percent occurrence scores for words
appearing in certain LIWC categories, which could then be weighted with
valence scores from Warriner et al. (2013). The researchers will
consider other applications of the LIWC in future studies aiming to
analyze political content. James Pennebaker (2007), one of the creators
of LIWC, has suggested the program could be applied to analyses of
cognitive, social, and emotional states. Furthermore, there is some
evidence LIWC can be used to detect deception, but the reliability of
the program's accuracy in this regard (67\%) calls this assertion into
question (Pennebaker, 2011). Despite the promising extent of LIWC's
abilities, without a firm theoretical foundation for such analyses, any
such applications of LIWC would face serious questions as to their
relevance, validity, and generalizability. Nevertheless, LIWC analyses
may represent a path to further research into political differences in
news media as well as theory formation.

There is, however, another theoretical framework that could prove useful
when applied to the method established in the present research.
Relational Frame Theory originates in applied behavior analysis, but
there exists recent scholarship that demonstrates its potential
usefulness in the political realm. Relational Frame Theory (RFT) holds
that stimuli, including speech and discourse, elicit responses in terms
of relationships to other stimuli, even when the stimuli do not
otherwise appear to be related (Blackledge, 2014). For example, an
individual's equivalent fear response to a venomous snake as well as a
wooded area said to \enquote{contain snakes} represents \enquote{mutual
entailment.} That is, the fear response is an effect of both the snake
and environment stimuli, which cause the fear response.
\enquote{Combinatorial entailment} can be represented by a fear response
elicited by a novel stimulus whose relationship to a well-known stimulus
elicits a certain response (Blackledge, 2014). In the snake example, if
another stimulus (Stimulus A), were said to be \enquote{worse than} the
snake, the relationship between A and the snake would elicit a more
severe fear response to Stimulus A. RFT utilizes a set of
\enquote{families} describing the nature of these relationships. Such
families include \enquote{coordinative,} establishing similarities
between stimuli, and \enquote{distinctive,} implying fundamental
differences between stimuli (Belisle, Paliliunas, Dixon, \& Tarbox,
2018). These families can be used to analyze networks of relational
frames in discourse.

Belisle et al. (2018) conducted a text analysis (using a computer
algorithm) of several speeches made by four recent U.S. presidents under
Relational Frame Theory. These speeches, representing prominent examples
of political discourse, were found to differ in their use of relational
frames. For example, speeches by Donald Trump were found to contain more
distinctive frames than other presidents, while Barack Obama utilized
more coordinative frames than his counterparts (Belisle et al., 2018).
At first glance, the differences appear to fall along party lines,
though other factors may be involved as well. In any case, this research
appears to indicate that RFT might be useful for analyzing political
news content similar to the method employed in the current study.

>>>>

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

@@

\abstract{
The media ecosystem has grown, and political opinions have diverged such
that there are competing conceptions of objective truth. Commentators
often point to political biases in news coverage as a catalyst for this
political divide. The Moral Foundations Dictionary (MFD) facilitates
identification of ideological leanings in text through frequency of the
occurrence of certain words. Through web scraping, the researchers
extracted articles from popular news sources' websites, calculated MFD
word frequencies, and identified words' respective valences. This
process attempts to uncover news outlets' positive or negative
endorsements of certain moral dimensions concomitant with a particular
ideology. In Experiment 1, the researchers gathered poltical articles
from four sources. We were unable to reveal significant differences in
moral or political endorsements, but we solidified the method to be
employed in further research. In Experiment 2, the researchers expanded
their number of sources to 10 and analyzed articles that pertain to two
specific topics: the 2018 confirmation hearings of U.S. Supreme Court
Justice Brett Kavanaugh and the partial U.S. Goverment Shutdown of
2018-2019. Once again, no significant differences in moral or political
endorsements were found.

@@

Moral Foundations Theory has received criticism on the grounds that its
assumptions regarding moral intuitions have little empirical basis.
Suhler, Churchland, and Joseph (2011) list several potential weaknesses
of MFT that they argue might threaten the theory's validity. First, the
authors challenge Haidt and Graham (2007)'s claims the moral intuitions
represented by MFT are innate and modular. Suhler et al. (2011) claim
that advances in biological sciences (embryology and microbiology,
specifically) make it more difficult for researchers to claim any one
trait is either innate or learned through experience. Rather, behaviors
likely result from interactions between genetics and experience (Suhler
et al., 2011). According to the authors, without solid data supporting
the innateness of moral foundations, Haidt and Graham have little from
which to make such a claim. Similarly, Haidt and Graham (2007) rely on
evidence authored by evolutionary psychologists to make a
\enquote{strong modularity claim} (Suhler et al., 2011). However, as
with innateness, there is little neurobiological evidence to support
modularity.

Suhler et al. (2011) also criticized the content and taxonomy of the
five foundations. The authors criticize Haidt and Graham (2007)'s
omissions of additional foundations, including \emph{industry} and
\emph{modesty}, claiming these concepts are moralized in many societies
worldwide. Likewise, the authors question whether or not the foundations
are sufficiently distinct as to stand as their own foundation. For
example, Suhler et al. (2011) posit that \emph{ingroup/loyalty} is
merely a group-focused version of \emph{harm/care}. Work by Graham et
al. (2011) might serve to rebut this criticism, as the researchers found
their original five-factor structure seemed to best fit the data when
validating the Moral Foundations Questionnaire. Finally, Suhler et al.
(2011) point out that particular concepts related to a foundation,
including \enquote{anger} in \emph{fairness/reciprocity} and
\enquote{deception} in \emph{ingroup/loyalty}, could be ascribed to any
of the other foundations as well. In other words, it becomes difficult
to recognize a particular concept as indicative of any one foundation
when, in theory, they could be applied to all five (Haidt \& Graham,
2007; Suhler et al., 2011).

These criticisms of Moral Foundations Theory are valid and should be
taken into consideration when conducting research with instruments
derived from MFT. Such criticisms are especially valid when considered
alongside questions about the Moral Foundations Dictionary in
particular. Nevertheless, there exists compelling evidence supporting
the validity of Moral Foundations Theory, albeit regarding its
application solely through the Moral Foundations Questionnaire. This
evidence is discussed within a brief explanation of the Moral
Foundations Questionnaire and its relationship to the Moral Foundations
Dictionary in the next section.

@@

In order to capture language's role in moral and political reasoning,
Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD)
in order to capture moral reasoning and justification as used in speech
and text. The MFD is composed of 259 words, with around 50 words
assigned to each of the five foundations. Graham et al. (2009) created a
preliminary list of words that they believed would be associated with
the five foundations. Then, using the Linguistic Inquiry and Word Count
(LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they
analyzed transcripts of liberal and conservative Christian sermons in
order to obtain frequencies of the occurrence of words from the
researchers' initial list. Graham et al. (2009) manually checked the
results from LIWC in order to make sure the results make sense given the
contexts and rhetorical devices used in the sermons, as word frequency
analysis ignores sentence context. The researchers offered the following
example from a Unitarian sermon as a demonstration of ambiguous
statements requiring human verification: \enquote{Don't let some
self-interested ecclesiastical or government authority tell you what to
believe, but read the Bible with your own eyes and open your heart
directly to Jesus} (Graham et al., 2009). This sentence added to the
\emph{authority/respect} total in LIWC's analysis, but it appears to
suggest that one should reject authority in this context. The
researchers eliminated this sentence from the \emph{authority/respect}
raw count on account of this discrepancy between the use of
authority-related words and the speaker's clear intentions (Graham et
al., 2009). The original Moral Foundations Dictionary can be found in
the Appendix.

@@

The Moral Foundations Questionnaire has been validated by multiple
researchers. Likewise, its five-factor structure has been demonstrated
to fit data in multiple countries (Davies, Sibley, \& Liu, 2014; Graham
et al., 2011). In their article introducing the 30-item MFQ, Graham et
al. (2011) conducted an exploratory factor analysis (EFA) and found that
a five-factor model fit the data better than a one, two, or three-factor
model. Davies et al. (2014) conducted a confirmatory factor analysis
(CFA) of the MFQ with a sample from New Zealand and likewise found the
five-factor model provided the best fit. While Davies et al. (2014)
concede the US and New Zealand share many similarities as Western
nations, which could raise questions regarding the validity of the MFQ
in non-Western nations. However, there are striking differences between
the two countries, including the lack of a two-party political system in
New Zealand, that provide grounds for claiming the MFQ generalizes
beyond the United States. Furthermore, Graham et al. (2011) claim to
find a \enquote{reasonable} degree of generalizability for the MFQ
across participants from many different regions in the world. These two
bodies of work also provide the best available evidence that the five
moral foundations are sufficiently distinct from one another, though
broader criticisms of MFT raised by Suhler et al. (2011) should still be
taken into account in studies involving the theory.

@@

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
sources (identified using CSS language) and compiled them into a dataset
(Wickham, 2016). Using this dataset, we identified word count and
average word count per source. This process was completed once daily
starting in February 2018 until March 2018. Starting in mid-March 2018,
the process was completed twice daily - once in the morning and again in
the evening. Data collection was terminated once 250,000 words per
source was collected in April 2018.

@@

As seen in Table \ref{tab:exp1-source-descriptives}, the sources are
similar in some aspects yet different in others. Valence appears to be
slightly positive across all sources. The large standard deviations seem
to indicate little to no presence of a difference in valence across
sources. \emph{The New York Times} published the greatest number of
articles as well as total words. \emph{Breitbart} featured the lowest
number of articles, and \emph{NPR} the lowest number of total words from
all articles. Per individual article, however, \emph{Breitbart} appears
to feature the highest average number of words as well as unique words.
Once again the standard deviations call into question any apparent
differences between sources. Finally, \emph{Fox News} articles had the
lowest reading grade level on average, while \emph{The New York Times}
ihad the highest. This result might be attributable to the greater
number of tokens in the average \emph{New York Times} article compared
to \emph{Fox News}. The standard deviations for readability suggest the
presence of a diverse array of articles for each source, ranging from
low to high reading level. Large standard deviations suggest the sources
feature a lot of overlap between them in their representation of scores.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp1-table}Experiment 1 Results - Multilevel Model}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .850 & 0.01\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .715 & 0.02\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .789 & -0.04\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.17 & .087 & -0.20\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.37 & .141 & -0.09\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

@@

Table \ref{tab:exp2-source-descriptives-kav} displays the descriptive
statistics for sources' writing on the Kavanaugh confirmation hearing.
The sources were similar in most basic linguistic aspects, except for
number of articles. For example, \emph{Sean Hannity} appears to have
published only 27 articles while \emph{Breitbart} published 757 articles
on this topic. Valence was found to be slightly positive across all
sources. \emph{Fox News} produced the most total words with the most
tokens on average. This is likely due to the fact \emph{Fox News}
transcribes many of their videos and publishes them in article form.
\emph{Politico} featured the highest number of types on average.
\emph{Rush Limbaugh} featured the lowest readability score on average by
grade level while \emph{Slate} featured the highest grade-level
readability score. The large standard deviations for these statistics,
however, preclude conclusions regarding differences in the sources' use
of language, as there is likely a lot of overlap between sources' use of
language.

Table \ref{tab:exp2-source-descriptives-gs} displays descriptive
statistics for articles about the partial government shutdown. Like the
Kavanaugh hearing, the sources were similar in average valence (slightly
positive). Once again, there was variation in the number of articles
published by each source on this topic. \emph{Sean Hannity}, \emph{Rush
Limbaugh}, and \emph{The Blaze} published fewer than 100 articles while
\emph{Fox News} published 1,013 articles. \emph{Fox News} again featured
the most total words and mean tokens, but this is likely due to to the
presence of a large amount of video transcriptions that the organization
published as articles. \emph{Politico} had the most types on average.
For this topic, \emph{Fox News} featured the lowest reading grade level
while \emph{Slate} featured the highest reading grade level. For each
statistic, the excessively high standard deviations render any
assertions regarding linguistic differences inconclusive on a
descriptive level due to the aforementioned overlap in sources' language
use.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablekav}Kavanaugh Results - Multilevel Model}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.39 & 1.35 & 0.66 & 1.35 & 2.89 & .020 & -0.19\\
Fairness/Reciprocity & 1.26 & 1.27 & 1.06 & 1.27 & -1.58 & .154 & 0.16\\
Ingroup/Loyalty & 1.15 & 1.22 & 1.06 & 1.22 & -0.63 & .549 & 0.07\\
Authority/Respect & 0.85 & 0.99 & 0.75 & 0.99 & -0.93 & .380 & 0.10\\
Purity/Sanctity & 0.90 & 1.28 & 1.06 & 1.28 & 0.88 & .406 & -0.11\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablegs}Government Shutdown Results - Multilevel Model}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.91 & 1.27 & 1.01 & 1.27 & 1.68 & .132 & -0.07\\
Fairness/Reciprocity & 1.22 & 1.05 & 0.91 & 1.05 & -0.76 & .468 & 0.27\\
Ingroup/Loyalty & 1.38 & 1.16 & 1.11 & 1.16 & -0.48 & .644 & 0.21\\
Authority/Respect & 0.87 & 1.24 & 0.36 & 1.24 & -1.38 & .203 & 0.37\\
Purity/Sanctity & 1.29 & 1.13 & 0.99 & 1.13 & 0.03 & .979 & 0.22\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

@@

The results obtained in both experiments, while not confirmatory,
provide several avenues for future investigation. Twice confronted with
the failure to uncover differences in news media's political content,
the researchers were unable to confirm the existence of an effect of
political lean on American news writing under the auspices of Moral
Foundations Theory. This does not, however, rule out the possibility
that news media is biased. As Mitchell et al. (2014) has demonstrated,
self-reported political partisans are drawn to separate sources, but the
present research was unable to offer any content-based explanations of
this phenomenon. Even when attenuating the scraping criteria to collect
text concerning (what the researchers believed to be) morally resonant
topics, the method failed to detect differences in MFD word use by
prominent news sources. Given part of the researchers' goal with this
study was to answer a method question, they must contemplate whether or
not the method or the construct of interest (political lean) are the
primary contributor to null results. Either the method used to detect
lean was insufficent for this purpose, or political lean simply does not
exist, indicating the presence of a true negative in the results.

The researchers believe these results represent a false negative rather
than a true negative. That is, we believe political lean exists, but
that we have been unable to detect it at a content level. The
researchers again turn to Mitchell et al. (2014)\enquote{s work in this
area: political partisans prefer certain news outlets over others. There
must be something about the sources} content that facilitates partisans'
repeated usage. Otherwise, they might not express such loyalty to any
one source (or group of sources) over others. Given this aspect of the
news economy, the researchers still believe certain news outlets in the
U.S. contatin a political slant. The researchers' focus, therefore,
turns toward the limitations of the method, especially the chosen
theoretical framework, under which the present research was conducted.

@@

There remains an aspect of language which this method still cannot
address: context. Owing to the presence of millions of words in the
analyses presented in the current study, the researchers would be unable
to check for context and correct endorsement scores accordingly. Graham
et al. (2009) actually performed a process of human error-checking for
context when formulating the MFD. While this might have helped that team
fix mistakes made by the LIWC program (Pennebaker et al. (2007)), manual
coding of this nature has the potential to introduce an unnacceptable
level of bias in the method, as perceptions of context would be
dependent upon raters' interpretations and beleif systems. Additionally,
the numerous ways in which context affects meaning make it difficult to
interpret MFD word occurrence outside of a narrow operational definition
of \enquote{endorsement.} In fact, political lean may be communicated
through methods other than through the use of particular words.

One method through which news organizations might appeal to political
partisans is through their content strategy rather than their use of MFD
words. The results in the current study demonstrate a similar pattern of
endorsement across both conservative and liberal sources. Specifically,
the researchers found slightly positive endorsements for all five
foundations no matter the political lean of the source. While this may
simply be an indication of insufficient MFD word occurrence in news
articles, it may also demonstrate truly equal endorsement despite
political lean. While news organizations may endorse all moral
foundations, they may also adopt strategies for doing so that appeal to
the political partisans consuming their product. As an example, a
liberal source might endorse \emph{purity/sanctity} through an appeal to
reduce waste or pollution (i.e.~keeping the earth \enquote{clean}). A
conservative source, on the other hand, might endorse
\emph{purity/sanctity} through an argument in favor of abstinence or
religious commitment (i.e.~keeping a \enquote{clean} soul, mind, or
body). Both publications would register as positive endorsement of
\emph{purity/sanctity} despite appropriating the concept for distinct
purposes. Through these divergent strategies, liberal and conservative
news organizations might be able to appeal to their fans while
exhibiting similar levels of moral endorsement. This could represent a
further weakness of the Moral Foudnations Dictionary as an instrument,
as it is unable to identify these potential strategies for communicating
certain points of view.

Thus, the researchers' attention turned to the instrument used to detect
differences: the Moral Foundations Dictionary. As mentioned before, MFD
is a measure developed from Moral Foundations Theory (Graham et al.,
2009). Based upon the results obtained, it might be necessary to
investigate alternative instruments that could better elucidate the
differences of interest. Going beyond specific instruments, other
theoretical perspectives may be more equipped to explain political
differences in discourse. Likewise, an atheoretical approach in which
large quantities of data are collected from which theories are
formulated may be best suited to this area of research.

Alternative instruments could be developed through the Linguistic
Inquiry and Word Count program (LIWC) (Pennebaker et al., 2007). As
stated before, the Moral Foundations Dictionary was established through
an analysis of Christian sermons performed mostly with LIWC as well as
some human-enforced corrections due to context (Graham et al., 2009).
This approach was grounded in Moral Foundations Theory, a clear
framework within which the authors developed the MFD. The original
authors of the MFD established sound psychometric properties for the
tool, implying its validity in several applications. For this reason,
the researchers elected to utilize the MFD as their primary instrument.

Instruments that improve upon the Moral Foundations Dictionary could
theoretically be developed through a LIWC analysis. Researchers could
obtain percent occurrence scores for words appearing in certain LIWC
categories, which could then be weighted with valence scores from
Warriner et al. (2013). The researchers will consider other applications
of the LIWC in future studies aiming to analyze political content.
Pennebaker et al. (2007) has suggested the program could be applied to
analyses of cognitive, social, and emotional states. Furthermore, there
is some evidence LIWC can be used to detect deception, but the
reliability of the program's accuracy in this regard (67\%) calls this
assertion into question (Pennebaker, 2011). Despite the promising extent
of LIWC's abilities, without a firm theoretical foundation for such
analyses, any such applications of LIWC would face serious questions as
to their relevance, validity, and generalizability. Nevertheless, LIWC
analyses may represent a path to further research into political
differences in news media as well as theory formation.
